### 2023-05-03 Notes/main.tex
Recapped lohr2017combining.

% \input{estimating_response_mod.tex}

\section*{Minimizing the Variance}

\begin{itemize}
  \item The goal of this section is to find optimal values of $b_2(X, Y_1)$ and
    $a_2(X, Y_2)$ such that the variance of $\hat \theta_{eff}$ is minimized.

  \item Recall:

    \begin{align*}
      \hat \theta_{eff} - \hat \theta_n 
      &= n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] \left(1 - \frac{R_{1i}}{\pi_{1+}}
      - \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \\
      &+ n^{-1} \sum_{i = 1}^n b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
      &+ n^{-1} \sum_{i = 1}^n a_2(X_i, Y_{2i}) \left(\frac{R_{2i}}{\pi_{2+}} -
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)\\
      &+ n^{-1} \sum_{i = 1}^n g_i \left(\frac{R_{1i}R_{2i}}{\pi_{11}} -
      1\right) \\
      &\equiv A + B + C + D.
    \end{align*}

  \item Notice that we will suppress the fact that response models are functions
    of $X$ (i.e. we write $\pi_{11}$ instead of $\pi_{11}(X)$).

  \item To compute the variance, we first solve for each covariance combination.
    Basically, all these computations rely on the following ideas. First, we
    assume that the response model is correctly specified. Consequently,
    $E[A] = E[B] = E[C] = E[D] = 0$ and things work out better. This helps when
    we take the covariance conditional on $X$ because the inner expectations are
    zero. The second key insight it to notice that $E[R_{j}^k] = E[R_j]$ for $j
    \in \{1, 2\}$ and $k \in \mathbb{N}$. This is because $R$ is a binary
    variable. Third, since we assume that the response models are correctly
    specified, we have $E[R_1 \mid X] = \pi_{1+}$, $E[R_2 \mid X] = \pi_{2+}$,
    and $E[R_1, R_2 \mid X] = \pi_{11}$.

    The overall approach to each of these computations is the following:
    (1) take conditional expectations with respect to $X$ (the $\Cov(E[\cdot])$
    term is zero), (2) expand the covariance to $E[XY] - E[X]E[Y]$ (the second
    term is also zero), (3) by the MAR assumption $g, a_2, b_2$ are independent
    of $R_1$ and $R_2$ and we can take the latter out of the expectation, (4)
    evaluate and simplify expressions involving $E[R]$.

    \begin{align*}
      \Cov(A, B) &= n^{-2} \sum_{i = 1}^n E\left[\Cov\left(E[g_i \mid X] \left(1
      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X_i,\right. \right.\\
      &\qquad \qquad \qquad \qquad b_2(X_i, Y_{1i}) \left. \left.
      \left(\frac{R_{1i}}{\pi_{1+}} - \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid
      X_i\right)\right]\\
      &= n^{-1} E\left[E\left[E[g \mid X] \left(1 - \frac{R_{1i}}{\pi_{1+}} -
      \frac{R_{2i}}{\pi_{2+}} + \frac{R_{1i}R_{2i}}{\pi_{11}}\right)
      b_2(X_i, Y_{1i}) \left(\frac{R_{1i}}{\pi_{1+}} -
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) \mid X \right]\right] \\
      &= n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
    \end{align*}

    By symmetry,
    \begin{align*}
      \Cov(A, C) 
      &= n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(A, D) &= n^{-1} E\left[E\left[E[g \mid X] \left(1 -
      \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right) g \left(\frac{R_1 R_2}{\pi_{11}} -
      1\right) \mid X\right] \right]\\
      &= n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
      \frac{1}{\pi_{2+}} + 2\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(B, C) &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
      \left(\frac{R_2}{\pi_{2+}} - \frac{R_1 R_2}{\pi_{11}}\right) \mid
      X\right]\right]\\
      &= n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(B, D) &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
      E\left[\left(\frac{R_1}{\pi_{1+}} - \frac{R_1 R_2}{\pi_{11}}\right)
      \left(\frac{R_1 R_2}{\pi_{11}} - 1\right) \mid X\right] \right]\\
      &= n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    By symmetry,
    \begin{align*}
      \Cov(C, D) 
      &= n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    We also compute the variance terms,

    \begin{align*}
      \Cov(A, A) &= n^{-1} E\left[ E\left[ E[g \mid X]^2 \left(1
      - \frac{R_{1i}}{\pi_{1+}} - \frac{R_{2i}}{\pi_{2+}} +
      \frac{R_{1i}R_{2i}}{\pi_{11}}\right)^2 \mid X\right] \right]\\
      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(B, B) &= n^{-1} E\left[ E\left[b(X, Y_1)^2 \left(\frac{R_1}{\pi_{1+}}
      - \frac{R_1 R_2}{\pi_{11}}\right)^2 \mid X\right] \right]\\
      &= n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
      \frac{1}{\pi_{11}}\right)\right].
    \end{align*}

    \begin{align*}
      \Cov(C, C) &= n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right].
    \end{align*}

    \begin{align*}
      \Cov(D, D) &= n^{-1} E\left[ E\left[g_i^2 \left(\frac{R_1 R_2}{\pi_{11}} -
      1\right)^2 \mid X\right] \right]\\
      &= n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
      1\right)\right].
    \end{align*}

    This means that
    \begin{align*}
      \Var(\hat \theta_{eff} - \hat \theta_n) 
      &= \Cov(A, A) + 2 \Cov(A, B) + 2 \Cov(A, C) + 2 \Cov(A, D) + \Cov(B, B) \\
      &+ 2 \Cov(B, C) + 2 \Cov(B, D) + \Cov(C, C) + 2 \Cov(C, D) + \Cov(D, D) \\
      &= n^{-1} E \left[ E[g \mid X]^2 \left(-1 + \frac{2 \pi_{11}}{\pi_{1+}
      \pi_{2+}} - \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right] \\
      &+ 2 n^{-1} E\left[E[g \mid X] E[b_2(X, Y_1) \mid X]
      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
      &+ 2 n^{-1} E\left[E[g \mid X] E[a_2(X, Y_2) \mid X]
      \left(\frac{1}{\pi_{1+}} + \frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}} -
      \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right]\\
      &+ 2 n^{-1} E\left[E[g \mid X]^2 \left(\frac{-1}{\pi_{1+}} -
      \frac{1}{\pi_{2+}} + 2\right)\right]\\
      &+ n^{-1} E\left[ E[b_2(X, Y_1)^2 \mid X] \left(\frac{-1}{\pi_{1+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &+ 2 n^{-1} E\left[E[b_2(X, Y_1) \mid X] E[a_2(X, Y_2) \mid X]
      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - \frac{1}{\pi_{1+}} -
      \frac{1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right)\right]\\
      &+ 2 n^{-1} E\left[ E[b_2(X, Y_1) \mid X] E[g \mid X]
      \left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{11}}\right)\right]\\
      &+  n^{-1} E\left[ E[a_2(X, Y_2)^2 \mid X]
      \left(\frac{-1}{\pi_{2+}} + \frac{1}{\pi_{11}}\right) \right]\\
      &+ 2 n^{-1} E\left[ E[a_2(X, Y_2) \mid X] E[g \mid X]
      \left(\frac{1}{\pi_{2+}} - \frac{1}{\pi_{11}}\right)\right]\\
      &+ n^{-1} E\left[ E[g^2 \mid X] \left(\frac{1}{\pi_{11}} -
      1\right)\right].
    \end{align*}

    Differentiating yields:

    \begin{align*}
      \frac{\partial}{\partial a_2} \Var(\hat \theta_{eff} - \hat \theta_n)
      &= E\left[E[g \mid X] \left(\frac{1}{\pi_{1+}} + \frac{2}{\pi_{2+}} -
      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right)\right] \\
      &+ E\left[E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &+ E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &\equiv 0, \text{ and }\\
      \frac{\partial}{\partial b_2} \Var(\hat \theta_{eff} - \hat \theta_n)
      &= E\left[E[g \mid X] \left(\frac{2}{\pi_{1+}} + \frac{1}{\pi_{2+}} -
      \frac{2}{\pi_{11}} - \frac{\pi_{11}}{\pi_{1+}\pi_{2+}}\right)\right]\\
      &+ E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
      \frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &+ E\left[ E[b_2(X, Y_2) \mid X] \left(\frac{-1}{\pi_{1+}} +
      \frac{1}{\pi_{11}}\right)\right]\\
      &\equiv 0. 
    \end{align*}

    Substitution shows that these constraints are equivalent to:

    \begin{align*}
      E&\left[E[g \mid X] \left(\frac{-1}{\pi_{1+}} +
    \frac{1}{\pi_{2+}}\right)\right] + E\left[ E[b_2(X, Y_1) \mid X]
    \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -\frac{1}{\pi_{2+}}\right)\right]\\
      &- E\left[E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} - 
    \frac{1}{\pi_{1+}}\right)\right] \equiv 0
    \end{align*}

    where is the same as,

    \begin{align*}
      E&\left[ E[b_2(X, Y_1) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
    \frac{1}{\pi_{2+}}\right) \right] + E\left[E[g \mid X]
      \left(\frac{1}{\pi_{2+}}\right)\right] \\
      &=
    E\left[ E[a_2(X, Y_2) \mid X] \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}} -
    \frac{1}{\pi_{1+}}\right) \right] + E\left[E[g \mid X]
    \left(\frac{1}{\pi_{1+}}\right)\right].
    \end{align*}

  \item These constraints can be satisfied (this is sufficient but maybe not
    necessary) if

    \begin{align*}
      E&\left[(E[b_2(X, Y_1) \mid X] - E[a_2(X, Y_2) \mid X])
      \left(\frac{\pi_{11}}{\pi_{1+} \pi_{2+}}\right) \right] = 0\\
      E&\left[\left(\frac{1}{\pi_{1+}} - \frac{1}{\pi_{2+}}\right)(E[a_2(X, Y_2)
      \mid X] + E[b_2(X, Y_1) \mid X] - 2E[g \mid X])\right] = 0.
    \end{align*}

%\textbf{Questions for Dr. Kim}
%\begin{itemize}
%  \item Where do we go from here?
%  \item This is a functional expectation that I need to calibrate. Should I use
%    some sort of basis spline to try to estimate these expectations and then set
%    them equal to zero? I don't think that I have worked with this kind of
%    constraint before.
%\end{itemize}

@@

@@

>>>>

%\input{estimating_response_mod.tex}

@@

\input{minimizing_variance.tex}

@@

\input{multiphase_estimators.tex}
